Certified Knowledge Compilation \\ with Application to Verified Model Counting

### 2023-04-28 doc/sat2023-rebuttal.txt

We wish to thank the reviewers for their helpful comments.  We will first respond to the most important items overall and then to ones by specific reviewers

DETERMINISTIC vs. DECISION
We will be more careful about the distinctions between these two flavors of DNNFs.  Our proof framework supports DETERMINISTIC, although our current proof generator relies on the inputs being in DECISION form.  We will clarify this.

PRIOR WORK
It was a major oversight on our part not to have cited the work by Capelli et al. AAAI-21.  This is, of course, very relevant.  Looking in detail at the conditions they check, it can be seen that we do much of the same work in our proof generator.  The main difference is that we then emit information as lines in a well-defined, formally-verified clausal proof framework.

We have been unable to locate a copy of their checker online.  If we can find one, we’ll do a comparison of runtimes and proof sizes.  D4 can emit a DRAT version of its conflict clauses.  At the very least, we will measure the sizes and checking times for these.

GENERALITY
Although we present results only for results from D4, we have also successfully used it on the results from DSHARP.  We have found that DSHARP lags far behind D4 in terms of both time and the size of the generated result.  Our proof framework is sufficiently general to handle SDDs.  We will list that as future work.

Your reviews sparked us to consider whether we could deal with formulas that have been preprocessed using equivalence-preserving transformations.  We were able to modify the source code of D4 to get it to perform preprocessing (it was disabled) and are collecting data now.  The structured approach does not work, but the monolithic approach works fine.  We will present data on this.  We do not believe the CD4 program is able to generate end-to-end proofs of this form.

REVIEWER#1

Your suggestions on how to make the presentation more clear are helpful.  The flow is not ideal, in part because we had to move so much information to the appendix to stay within the page limit

REVIEWER#2

Thank you for notifying us about prior work on d-Ds.  These are exactly what we mean by POGs, and we will include citations.

Regarding a proof of correctness.  We will work to make it easier to read the portions of the Lean presentation.  We feel that the value of a formal proof in Lean far exceeds what we could express with an informal proof.

Regarding the question of whether the SAT solver could introduce extension variables.  We do not allow this, for exactly the reason you mention.  The only extension variables are those encoding the AND and OR operations.

Your observation about the reverse implication proofs being tractable is noted.  We see from the Capelli paper that Darwiche had already observed this, and we will cite that.  For a practitioner, we feel that a one-sided proof would be better than nothing.

### 2023-04-30 doc/mc-data.tex
Better visualization of clause data
Quantify effect of preprocessing

\input{figure-clauses-d4-verify}

@@

We conducted a set of experiments using the 80 benchmark instances for
which the total number of proof clauses when both optimizations are
performed was limited to $10^7$.  We then made three additional runs
of the CPOG generator and checker, with each of the two optimizations
disabled individually and in combination.  Figures
\ref{fig:optimized:partially-optimized}--\ref{fig:optimized:unoptimized} illustrates the effect of the two
optimizations.  Each set of points shows the number of clauses when
both optimizations are performed for the $X$ value and the number with
one or both of these optimizations disabled as the $Y$ value.

@@

Ran all benchmarks for which generation + checking could complete in 1000 seconds.  Total = 82.  All completed in less than 1000 seconds with structured.
Compare both runtimes and total number of clauses for structured vs. monolithic:
\begin{itemize}
\item Structured complete 83.
\item Monolithic completed  70,timed out on 13.
\item Of the 70 completed, 23 were faster with monolithic, 37 were slower (min ratio = 0.015, max ratio = 9.66, hmean = 0.40)
\item Of the 70 completed, 60 had fewer total clauses with monolithic, 10 had more (min ratio = 0.07, max ratio = 1.09, hmean = 0.49)
\end{itemize}

>>>>

\input{figure-clauses-d4-result}

@@

We conducted a set of experiments using the 80 benchmark instances for
which the total number of proof clauses when both optimizations are
performed was limited to $10^7$.  We then made three additional runs
of the CPOG generator and checker, with each of the two optimizations
disabled individually and in combination.  Figures
\ref{fig:optimized:partially-optimized}--\ref{fig:optimized:lessoptimized} illustrates the effect of the two
optimizations.  Each set of points shows the number of clauses when
both optimizations are performed for the $X$ value and the number with
one or both of these optimizations disabled as the $Y$ value.

@@

Ran all benchmarks for which generation + checking could complete in 1000 seconds.  Total = 83.  All completed in less than 1000 seconds with structured.
Compare both runtimes and total number of clauses for structured vs. monolithic:
\begin{itemize}
\item Structured complete 83.
\item Monolithic completed  69,timed out on 14.
\item Of the 69 completed, 19 were faster with monolithic, 50 were slower (min ratio = 0.013, max ratio = 18.97, hmean = 0.41)
\item Of the 69 completed, 59 had fewer total clauses with monolithic, 10 had more (min ratio = 0.07, max ratio = 1.09, hmean = 0.49)
\end{itemize}

\section{Preprocessing}

\input{figure-runtime-preprocess}
\input{figure-clauses-preprocess}

Use preprocessor in D4.  Enable its three preprocessing methods
(backbone, vivification, occElimination), all of which preserve
equivalence.  Cannot use structured mode for forward implication
proof, since modified clauses do not decompose in the way anticipated.

Run same 83 benchmarks used to compare monolithic vs. structured mode.
Incorporate time to run D4 into evaluation, since differs according to
how D4 runs.

Runtime results:
\begin{itemize}
\item Monolithic without preprocessing completes 68  of then with an overall time less than 1000 seconds.  (One less than when only consider generation and checking).
\item Monolithic with prepreprocessing completes these plus one more (not the same as previously excluded)
\item Of the 68, 53 run faster with preprocessing and 15 run slower (mininum ratio 0.04, maximum ratio 1.65, harmonic mean 0.38)
\item Shifts only slightly when don't count D4 time.  Both complete 69 benchmarks. 56 faster, 13 slower (minimum ratio 0.04, maximum ratio 1.49, harmonic mean 0.37)
\end{itemize}

Clause results:
\begin{itemize}
\item Consider the 68 cases that completed for both with overall time less than 1000 seconds.
\item 27 had more clauses without preprocessing, 19 were identical, and 22 had less.
\item Min = 0.43, Max = 4.57, Harmonic mean = 0.98
\end{itemize}
